### 2019-04-30 report/paper.tex
Worked on the paper

%% wlkr
\documentclass[12pt,journal,compsoc]{IEEEtran}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage{float}
\usepackage{listings}

\begin{document}
\title{Final Project}

\author{Ryan Walker, Jin Seo}

\IEEEtitleabstractindextext{%
% \begin{abstract}
% This paper will outline events of the failed Ariane 5 maiden launch. In addition it will describe the failure mode and recommend alternative system design and practices that would have prevented the crash.
% \end{abstract}
}

% make the title area
\maketitle
\IEEEpeerreviewmaketitle

\section{Introduction}
\IEEEPARstart{R}{ecent} advancements in drones have made brushless DC motors (BLDC motors) very inexpensive and lightweight. This has sparked vivid open source development of various motor drivers [\ref{odrive}] and mechanical support components [\ref{opentorque}]. By leveraging these tools we are aiming to build a lightweight 6-axis compliant robot manipulator. In the essence of time, only 3 axis' will be developed and dummy mass will be placed on the end. The following metrics define success for the device.

\begin{center}
\begin{tabular}{ c|c } 
 Spec & Quantity  \\ 
 \hline
 Payload & 1kg \\ 
 Working Area & ? \\ 
 DOF & 3 \\ 
 Speed & ?  \\ 
 Cost & sub \$3k  \\ 
\end{tabular}
\end{center}

\subsection{Motion Control}
\label{odrive}
For the motion control we want something that is open source, can deliver high current and is closed loop. The ODrive ticks all these boxes and more. Thanks to the hard work of developers across the world we have the tools to design powerful electromechancs the can be used to solve real problems. The Odrive is capable of driving two motors and receiving two channels of encoder feedback. We will need to use 3 boards for the entire robot. This comes to a total cost of only 380\$.

\begin{figure}[H]
\centering
\caption{ODrive}
\includegraphics[scale=0.15]{images/odrive.png}
\end{figure}

The firmware is completely open source and will need to be modifies as we require CAN for communication.

\subsection{Motors}
The ubiquity of drones has reduced the cost of motors significantly. The Multistar 9235-100KV has been measured to deliver 4.7Nm \cite{9235T} without a gearbox. It comes at a mere cost of 89\$.

\begin{figure}[H]
\centering
\caption{9235-100KV}
\includegraphics[scale=0.55]{images/motor.jpg}
\end{figure}

\subsection{Gearbox}
OpenTorque is a powerful, compliant actuator for legged robotics. We planning on using it for our revolute robot arm. As each linkage in this style of robot takes away from the payload of the robot, 3d printing something like this out of nylon comes with a serious weight advantage.

\begin{figure}[H]
\centering
\caption{OpenTorque}
\includegraphics[scale=0.15]{images/openTorque.png}
\end{figure}

% Bib
\begin{thebibliography}{1}
\bibitem{9235T}
Torque measurements for various motors. 
\url{https://docs.google.com/spreadsheets/d/12vzz7XVEK6YNIOqH0jAz51F5VUpc-lJEs3mmkWP1H4Y/edit#gid=0}

\end{thebibliography}

\end{document}

The SRI has been operating on the Ariane 4 without failure for several launches. However, the horizontal velocity of the Ariane 5 is much higher than that of the Ariane 4. This caused an unexpectedly high value of an internal alignment function (BH routine), leading to a software exception. This exception was due to a failed number conversion: The code did an unprotected number conversion from a 64bit float to a 16bit signed integer. Since the float had a greater value than what could fit into the int, this caused a fatal error and the loss of 0.5 billion dollars. \\

The result of the BH function was not needed after liftoff, however the function continued to run for 40 seconds after liftoff occurred. This feature is based on a requirement for the Ariane 4, but is not necessary for the Ariane 5.\\

During the analyses of the software running on the embedded computer, it was found that seven variables were at risk of overflow \cite{501}. However, only four of these variables were protected. This was done because the unprotected variables were believed to be at either a physical limit or have a large margin of safety, which was clearly not the case. In addition a maximum workload of 80\% was set for the SRI computer.

\section{Fault Prevention}
% Burns Page 106, should this be Burns? Or Anderson and Lee?
Burns \cite{BURNS} stated that fault prevention is attempting to eliminate bugs before the system goes live. This can be done using the compiler, software testing, hardware testing and simulations.

\subsection{Code Maintenance}
The rocket was not even using the BH routine during the disaster. It was legacy code that was left in from the Ariane 4. This should have been caught during the design review and removed. This sufferers from the ``if it ain't broken don't fix it'' fallacy. In software design there can be plenty of code that ``ain't broken'', however when the entire system changes you can't just assume that everything will be fine. The running of the BH routine is contradictory to the 80\% workload excuse listed in Section \ref{fail}, if the team was so strapped for resources then why did they have useless code running for 40 seconds after launch? If this code had been removed the failure would not have occurred.

\subsection{Data Types}
\label{SD}
Without being able to find any schematics of the system I'm making the assumption that there was a 16bit databus which connected the SRI to the onboard computer. The SRI did the float conversion and then put the 16bit number on to the data bus.\\

It's a common misconception that floating point numbers are more precise than fixed point integers. This is not the case, floating point number offer more range. Which was the root of the problem, the number was too big. There are many arguments around ``protecting the conversion''. Protection of the conversion would have prevented the exception, however you still need to \textbf{fit} the number into a 16bit space if you're planning on doing a calculation with it. Just maxing the number could cause havoc with the control system\\

If the programmers did not need the precision of a 16bit int, then they should have put in into a 16bit float. This offers a max number of $\pm65504$, rather then the $\pm32768$ a signed int can take. Doing so would give a $2\times$ margin of error over the int. In addition the conversion should have been protected.

\begin{figure}[H]
\centering
\caption{16bit Float, source: \url{https://en.wikipedia.org/wiki/Half-precision_floating-point_format}}
\label{16}
\includegraphics{16b.png}
\end{figure}

However this technique will take a hit on the precision of the number and requires some conversion time. Another option is to remove an order of magnitude from the 16bit int, this would give more range and less precision as well. Without requiring the extra conversion time. It's up to the designer to decide if these are acceptable trade offs. \\

\subsection{Testing}
\label{testing}
The 501 report \cite{501} lists "There is no evidence that any trajectory data were used to analyse the behaviour of the unprotected variables, and it is even more important to note that it was jointly agreed not to include the Ariane 5 trajectory data in the SRI requirements and specification." \\

This decision wasn't the cause of the failure, however testing like this would have prevented it. The last step in designing a Full Fault Tolerant system is extensive testing. The telemetry data for the rocket could have been numerically calculated and fed into the system. The bug would have made itself obvious and the disaster would have been prevented. Burns \cite{BURNS} calls this fault removal and is a major component of fault prevention.

\section{Fault Tolerance}
% Burns Page 106, should this be Burns? Or Anderson and Lee?
Burns \cite{BURNS} stated that fault tolerance is  the ability for your system to keep running in the presence of faults. This is very challenging in embedded system as there typically can't be any human interaction. However when the hardware inevitably fails there should be routines to handle this. \\

% The Ariane 5 would have to be ``Full Fault Tolerant'', meaning that in the event of a fault then ship needs to keep flying.

\subsection{Redundancy}
\label{Red}
Redundant hardware is used extensively in Full Fault Tolerant system. Redundancy provides alternative hardware in the event of a failure of the main hardware. Burns \cite{BURNS} wrote about Triple Modular Redundant (TMR) systems where three identical pieces of hardware are running in parallel. If the output of one is too far off then the others, it's result is masked out in the computation. This same idea can be extended to $N$ pieces of hardware\\

The Ariane 5 had two parallel SRI systems running. This means that in the event of a sensor or CPU hardware failure it was possible to switch out to the backup SRI. However the failure was not a hardware failure, it was a software failure. So both the main and redundant systems failed from the same software failure. Therefore redundancy did not save the Ariane 5.\\

Burns \cite{BURNS} spoke of the software equivalent to TMR, so called ``$N$-version'' programming. This is a method of software design which which requires $N$-groups of software developers to develop their own codebase, without interaction with each other. The software is then run in parallel and takes the same form as TMR. \\

However Burns listed three major issues with $N$-Version Programming:

\begin{enumerate}
\item Poor Initial Specification
\item Lack of Independence of Design Effort
\item Inadequate Budget
\end{enumerate}

Sections \ref{testing} and \ref{execpt} outlined some serious issues with the initial spec, an implementation of $N$-Version programming would have suffered. The control system for the Ariane 5 is a complex system, so it would be challenging to prevent people from sharing resources and ideas. Finally, would three independent development paths be affordable? Typically the most expensive resource of an embeded system is the development time, I don't think it would be reasonable to $3\times$ this.\\

$N-Version$ Programming is a novel approach to the problem and could have been used to avoid the disaster, however the feasibility is still a question. I would recommend approaching the problem

\subsection{Exception Handling}
\label{execpt}
The specification for the design stated that during an event of a software exception, the system should indicate the failure on the databus to be stored in EEPROM memory. This is followed by a processor shutdown. This technique seems reasonable in three situations:
\begin{enumerate}
\item Debugging environment.
\item System with ``N-Version Programming'' where the OBC would detect the exception and another system can take over. (See Section \ref{Red}).
\item System that catches the exception and can self correct. 
\end{enumerate}
This technique in not reasonable for a system that has no backup routine for exceptions. The OBC read the software exception bitstream and interpreted as data. As the BH routines wasn't even being used at this time of the crash, ignoring the software exception would have been a better response and would have saved the rocket.\\

A somewhat reasonable approach would be to have understood that there was an exception on the databus and extrapolate data. This would be an absolute fallback routine that is still very non-ideal, but better than what was implemented.

\subsection{Conversion Protection}
Baber \cite{SW} has spoken vividly about the conversion. I've written some code below to demonstrate a possible solution that would offer the protection.

\input{$HOME/wlkrUtils/TeX/codeCBW.tex} %For C
\lstinputlisting{../conv.c}
After examining the code the flaw becomes obvious, is the event of an overflow the max number will be presented. It's possible to make range and precision trade offs, which is discussed more rigorously in Section \ref{SD}. If it's not possible to make these trade offs then a simple handshaking routine could be implemented, this is shown in Section \ref{Nconv}\\

\subsection{Handshaking For Conversion}
It's entirely possible that the team had the processing overhead to implement a handshaking protocol, shown below. 
\label{Nconv}
\lstinputlisting{../convN.c}
This is using a special bit pattern to indicate to the embedded computer that the incoming data will not be a 16bit integer, but a 32 bit integer. In addition code should be implemented to ensure that the 64bit number isn't higher than what can fit in the 32bit number.\\

In the system many of the conversions were actually protected. However the team claimed that they were not all protected because of the processing overhead. I would argue that this is more overhead, but not by much. This would do one comparison and would not run the code inside the if statement unless there in in fact an overflow.

\section{Conclusion}
When developing embedded software the absolute first line of defense is well written code. Without a solid foundation there's no amount of testing and fault handling that can save the project. That being said, bugs are typically found and fixed during testing afterwords. With an extensive test procedure and enough time spent running the system through all scenarios it can become extremely stable. The last line of defense is run time software fault handling such as $N$-Version programming and smart use of exceptions.\\

The team that prepared the Ariane 5 for its maiden launch failed at all three of these software development techniques. There was poorly written code which made hardware assumptions. Following proper code maintenance techniques would have solved the issue from the beginning. The team should have started from the drawing board and if any code was to be recycled, it should have been properly studied and understood. This would have solved the compatibility issues between the Ariane 4 and Ariane 5. There were completely inadequate test procedures. A proper testing plan should have been implemented with mock telemetry data. If the code maintenance above didn't catch the bug, then testing would have. Finally the run time exception handling didn't make any sense. There should have been something reasonable implemented in the case of a software exception such as $N$-Version redundancy or data extrapolation.

% Bib
\begin{thebibliography}{1}
\bibitem{501}
Prof. J. L. LIONS, "Flight 501 Failure", Report by the Inquiry Board
Paris, 19 July 1996

\bibitem{SW}
Robert L. Baber, "The Ariane 5 explosion as seen by a software engineer", Risks-18.89, 12 March 1997 
\url{http://www.cas.mcmaster.ca/~baber/TechnicalReports/Ariane5/Ariane5.htm} [Accessed March 12, 2019]

\bibitem{BURNS}
A. Burns, A. Wellings. \textit{Real-Time Systems and Programming Languages}, Pearson Canada, 2009.

\end{thebibliography}

\end{document}

### 2019-06-22 report/paper.tex
Worked on the writeup

@@
 \section{Introduction}
-\IEEEPARstart{R}{ecent} advancements in drones have made brushless DC motors (BLDC motors) very inexpensive and lightweight.
-This has sparked vivid open source development of various motor drivers [\ref{odrive}] and mechanical support components [\ref{opentorque}].
-By leveraging these tools we are aiming to build a lightweight 6-axis compliant robot manipulator.
-In the essence of time, only 3 axis' will be developed and dummy mass will be placed on the end.
-The following metrics define success for the device.
+\IEEEPARstart{B}{iologists} have been aiming to understand bird migration for decades.
+Various custom hardware has been developed for doing this \cite{BirdTracking}, however is it costly an requires capture and release of the bird.
+We are proposing the design, construction and deployment of a device that resides in a stationary location.
+This device is equip with a microphone and microcontroller, and by using machine learning \cite{ML1} \cite{ML2} can detect and speciate various types of birds.
+The device will also be able to record various types of environmental data for cross-correlation with bird density in post processing.
+After the successful implementation we desire to make a time series bird density heat map.
+
+\section{Machine Learning on the Edge}
+Machine learning and AI have been popular topics in the last couple of year, especially the use of neural nets in classification algorithms.
+It has not been until recently that these applications have moved into edge compute, meaning processing has to be run on the device, rather than the cloud.
+We propose to design a NN (Neural Network) that can be run on a STM32 microcontroller.
+This is a rapidly expanding field and there are various software implementations \cite{CMSIS}.
+
+\section{Hardware}
+The hardware will need to be battery powered in order to collect data over extended periods of time.
+It should be inexpensive so it becomes feasible to deploy many devices and should be completely stand alone.
+In addition it should have a wiresless uplink/downlink to transmit data to and from a computer.
+
+\begin{thebibliography}{1}
+
+\bibitem{BirdTracking}
+Bird Tracking Hardware.
+\url{https://atstrack.com/animal-class/avian.aspx} [Accessed June 22, 2019]
+
+\bibitem{ML1}
+Stowell, Wood, Pamuła, Stylianou4, Glotin "Automatic acoustic detection of birds through deep learning: the first Bird Audio Detection challenge"
+\url{https://arxiv.org/pdf/1807.05812.pdf}
+
+\bibitem{ML2}
+Ilyas Potamitis, "Deep learning for detection of bird vocalisations"
+\url{https://arxiv.org/pdf/1609.08408.pdf}
+
+\bibitem{CMSIS}
+CMSIS NN Software Library
+https://www.keil.com/pack/doc/CMSIS/NN/html/index.html
+
+\end{thebibliography}
+\end{document}
 
@@
 \begin{thebibliography}{1}
-\bibitem{501}
-Prof. J. L. LIONS, "Flight 501 Failure", Report by the Inquiry Board
-Paris, 19 July 1996
+%\bibitem{501}
+%Prof. J. L. LIONS, "Flight 501 Failure", Report by the Inquiry Board
+%Paris, 19 July 1996
 
@@
 
-\bibitem{BURNS}
-A. Burns, A. Wellings. \textit{Real-Time Systems and Programming Languages}, Pearson Canada, 2009.
-
 \end{thebibliography}
-
 \end{document}

### 2019-06-27 report/paper.tex
Worked on the paper

@@
 We are proposing the design, construction and deployment of a device that resides in a stationary location.
-This device is equip with a microphone and microcontroller, and by using machine learning \cite{ML1} \cite{ML2} can detect and speciate various types of birds.
-The device will also be able to record various types of environmental data for cross-correlation with bird density in post processing.
-After the successful implementation we desire to make a time series bird density heat map.
+This device is equip with a microphone and microcontroller, and by using machine learning techniques \cite{ML1} \cite{ML2} can detect and speciate various types of birds.
+In addition the device will also be able to record various types of environmental data for cross-correlation with bird density in post processing.
+Following the successful deployment we desire to make a bird density heat map of various regions and to undestand how it changes over time.
 
 \section{Machine Learning on the Edge}
-Machine learning and AI have been popular topics in the last couple of year, especially the use of neural nets in classification algorithms.
-It has not been until recently that these applications have moved into edge compute, meaning processing has to be run on the device, rather than the cloud.
-We propose to design a NN (Neural Network) that can be run on a STM32 microcontroller.
-This is a rapidly expanding field and there are various software implementations \cite{CMSIS}.
+Machine learning and AI have been popular topics in the last couple of years, especially the use of neural nets in classification algorithms.
+It has not been until recently that these applications have moved into edge compute, meaning all the processing is run on the device as apposed to cloud.
+We propose to design a NN (Neural Network) that can be run on a STM32 microcontroller for classification of various birds.
+As this is such a rapidly expanding field there are various software implementations developed for embeded applications \cite{CMSIS} \cite{TF}.
 
 \section{Hardware}
-The hardware will need to be battery powered in order to collect data over extended periods of time.
+The hardware will need to be battery powered in order to collect data over extended periods of time. We are currently targeting the specifications below.
+
+\begin{center}
+\begin{tabular}{ c | c }
+Spec             & Metric\\ 
+\hline 
+Battery Life     & 1 week\\
+Compatable Birds & 10    \\
+Cost (QTD 100)   & $<\$30$  \\
+\end{tabular}
+\end{center}
+
 It should be inexpensive so it becomes feasible to deploy many devices and should be completely stand alone.
 In addition it should have a wiresless uplink/downlink to transmit data to and from a computer.
+
+\subsection{Microcontroller}
+In addition to having experience with the STM32 family of microcontrollers, TensorFlow Embedded \cite{TF} has extensive documentation for implementation on the STM32F746 \cite{STM}.
+This part has 1MB flash, 320k RAM and runs at 216Mhz.
+TensorFlow embedded is designed to run on 20k of flash, so we will have more than enough space for the NN and application code.
+
+\subsection{Microphone}
+The device will contain a microphone that will be used to record ambient noises. A suitable microphone has not yet been selected.
+
+\section{Software}
+%% Write something about software here.
+%%
+%%
+
+\section{Predicted Difficulties}
+%% Write something about Difficulties.
+%%
+%%
 
@@
 
+\bibitem{TF}
+TensorFlow Lite for Microcontrollers
+https://www.tensorflow.org/lite/microcontrollers/overview
+
+\bibitem{STM}
+ARM® Cortex®-M7 STM32F7 Microcontroller IC 32-Bit 216MHz 1MB (1M x 8) FLASH 100-LQFP (14x14)
+https://www.digikey.ca/product-detail/en/stmicroelectronics/STM32F746VGT6/497-15819-ND/5287178
+
+\bibitem{Bird}
+xeno-canto is a website dedicated to sharing bird sounds from all over the world.
+https://www.xeno-canto.org/
 \end{thebibliography}
+
 \end{document}

### 2019-06-28 report/paper.tex
Moved jin's file into main document. Added some extra fields.

@@
 \usepackage{listings}
+\usepackage{tikz}
+\usetikzlibrary{positioning}
 
 \begin{document}
-\title{Final Project}
-
+\title{Bird Density, with Edge ML}
 \author{Ryan Walker, Jin Seo}
@@
 
+\tikzset{%
+  every neuron/.style={
+    circle,
+    draw,
+    minimum size=1cm
+  },
+  neuron missing/.style={
+    draw=none, 
+    scale=4,
+    text height=0.333cm,
+    execute at begin node=\color{black}$\vdots$
+  },
+}
+
 \section{Introduction}
@@
 \section{Machine Learning on the Edge}
+\begin{tikzpicture}[x=1.3cm, y=1.3cm, >=stealth]
+
+\foreach \m/\l [count=\y] in {1,2,3,missing,4}
+  \node [every neuron/.try, neuron \m/.try] (input-\m) at (0,2.5-\y) {};
+
+\foreach \m [count=\y] in {1,missing,2}
+  \node [every neuron/.try, neuron \m/.try ] (hidden-\m) at (2,2-\y*1.25) {};
+
+\foreach \m [count=\y] in {1,missing,2}
+  \node [every neuron/.try, neuron \m/.try ] (output-\m) at (4,1.5-\y) {};
+
+\foreach \l [count=\i] in {1,2,3,n}
+  \draw [<-] (input-\i) -- ++(-1,0)
+    node [above, midway] {$I_\l$};
+
+\foreach \l [count=\i] in {1,n}
+  \draw [->] (output-\i) -- ++(1,0)
+    node [above, midway] {$B_\l$};
+
+\foreach \i in {1,...,4}
+  \foreach \j in {1,...,2}
+    \draw [->] (input-\i) -- (hidden-\j);
+
+\foreach \i in {1,...,2}
+  \foreach \j in {1,...,2}
+    \draw [->] (hidden-\i) -- (output-\j);
+
+\end{tikzpicture}
+
 Machine learning and AI have been popular topics in the last couple of years, especially the use of neural nets in classification algorithms.
@@
 
+\subsection{Architecture}
+\begin{itemize}
+\item{\textbf{MCU:} STM32F746.}
+\item{\textbf{SD card:} For recording data.}
+\item{\textbf{18650 Cell:} 2x 18654 Li-ion cells for power.}
+\item{\textbf{Microphone:} Microphone for recording sound.}
+\item{\textbf{Temperature Sensor:} Temperature sensor for recording ambient temperature.}
+\item{\textbf{Additional Sensors???:} Should we put anything else.}
+\end{itemize}
+
+\section{User Experience}
+Depending on the lifetime of the device, at a regular cadence the user will replace the batteries and SD card.
+As the device will then experience a power cycle, time will be lost.
+The device will then start recording from $t=0$.
+The user is required to keep track of the time of battery replacements.
+Real time is then implemented after the fact.
+
 \section{Software}
-%% Write something about software here.
-%%
-%%
+In the project we are planning on using Matlab/Simulink for any filter design, C/C++ for embedded code and possibly some python for post processing.
+
+\subsection{Signal processing}
+This project is required to utilize several signal processing techniques.
+Some of these techniques could be included filters, discrete Fourier transforms, and so on.
+Filters are necessary to improve the quality of the bird calls and remove unwanted noises.
+In addition, in order to come up with digitized data from acoustic analog signals, other techniques could be applied such as Discrete wavelet transform or mel-frequency cepstrum.
+In this case, MATLAB can be used for simulation.
 
 \section{Predicted Difficulties}
-%% Write something about Difficulties.
-%%
-%%
+We predict that it will be difficult to gather a a concise and useful dataset to train the neural network.
+However, there are sites that contain vast datasets \cite{Bird}.
+As every circumstance produces different audiable signals, finding proper filters to samples could be a major problem.
+Another major difficulty we predict is implementation of the  Neural Network and how we can apply it to our project, even as we are going to use a library, designing the neural structure for an embedded application might be challenging.
 
@@
 \end{document}
-
-\begin{center}
-\begin{tabular}{ c|c } 
- Spec & Quantity  \\ 
- \hline
- Payload & 1kg \\ 
- Working Area & ? \\ 
- DOF & 3 \\ 
- Speed & ?  \\ 
- Cost & sub \$3k  \\ 
-\end{tabular}
-\end{center}
-
-\subsection{Motion Control}
-\label{odrive}
-For the motion control we want something that is open source, can deliver high current and is closed loop.
-The ODrive ticks all these boxes and more.
-Thanks to the hard work of developers across the world we have the tools to design powerful electromechancs the can be used to solve real problems.
-The Odrive is capable of driving two motors and receiving two channels of encoder feedback.
-We will need to use 3 boards for the entire robot.
-This comes to a total cost of only 380\$.
-
-\begin{figure}[H]
-\centering
-\caption{ODrive}
-\includegraphics[scale=0.15]{images/odrive.png}
-\end{figure}
-
-The firmware is completely open source and will need to be modifies as we require CAN for communication.
-
-\subsection{Motors}
-The ubiquity of drones has reduced the cost of motors significantly. The Multistar 9235-100KV has been measured to deliver 4.7Nm \cite{9235T} without a gearbox. It comes at a mere cost of 89\$.
-
-\begin{figure}[H]
-\centering
-\caption{9235-100KV}
-\includegraphics[scale=0.55]{images/motor.jpg}
-\end{figure}
-
-\subsection{Gearbox}
-OpenTorque is a powerful, compliant actuator for legged robotics.
-We planning on using it for our revolute robot arm.
-As each linkage in this style of robot takes away from the payload of the robot, 3d printing something like this out of nylon comes with a serious weight advantage.
-
-\begin{figure}[H]
-\centering
-\caption{OpenTorque}
-\includegraphics[scale=0.15]{images/openTorque.png}
-\end{figure}
-
-% Bib
-\begin{thebibliography}{1}
-\bibitem{9235T}
-Torque measurements for various motors. 
-\url{https://docs.google.com/spreadsheets/d/12vzz7XVEK6YNIOqH0jAz51F5VUpc-lJEs3mmkWP1H4Y/edit#gid=0}
-
-\end{thebibliography}
-
-\end{document}
-
-The SRI has been operating on the Ariane 4 without failure for several launches.
-However, the horizontal velocity of the Ariane 5 is much higher than that of the Ariane 4.
-This caused an unexpectedly high value of an internal alignment function (BH routine), leading to a software exception.
-This exception was due to a failed number conversion: The code did an unprotected number conversion from a 64bit float to a 16bit signed integer.
-Since the float had a greater value than what could fit into the int, this caused a fatal error and the loss of 0.5 billion dollars.
-\\
-
-The result of the BH function was not needed after liftoff, however the function continued to run for 40 seconds after liftoff occurred.
-This feature is based on a requirement for the Ariane 4, but is not necessary for the Ariane 5.\\
-
-During the analyses of the software running on the embedded computer, it was found that seven variables were at risk of overflow \cite{501}.
-However, only four of these variables were protected.
-This was done because the unprotected variables were believed to be at either a physical limit or have a large margin of safety, which was clearly not the case.
-In addition a maximum workload of 80\% was set for the SRI computer.
-
-\section{Fault Prevention}
-% Burns Page 106, should this be Burns? Or Anderson and Lee?
-Burns \cite{BURNS} stated that fault prevention is attempting to eliminate bugs before the system goes live. This can be done using the compiler, software testing, hardware testing and simulations.
-
-\subsection{Code Maintenance}
-The rocket was not even using the BH routine during the disaster.
-It was legacy code that was left in from the Ariane 4.
-This should have been caught during the design review and removed.
-This sufferers from the ``if it ain't broken don't fix it'' fallacy.
-In software design there can be plenty of code that ``ain't broken'', however when the entire system changes you can't just assume that everything will be fine.
-The running of the BH routine is contradictory to the 80\% workload excuse listed in Section \ref{fail}, if the team was so strapped for resources then why did they have useless code running for 40 seconds after launch?
-If this code had been removed the failure would not have occurred.
-
-\subsection{Data Types}
-\label{SD}
-Without being able to find any schematics of the system I'm making the assumption that there was a 16bit databus which connected the SRI to the onboard computer.
-The SRI did the float conversion and then put the 16bit number on to the data bus.\\
-
-It's a common misconception that floating point numbers are more precise than fixed point integers.
-This is not the case, floating point number offer more range.
-Which was the root of the problem, the number was too big.
-There are many arguments around ``protecting the conversion''.
-Protection of the conversion would have prevented the exception, however you still need to \textbf{fit} the number into a 16bit space if you're planning on doing a calculation with it.
-Just maxing the number could cause havoc with the control system\\
-
-If the programmers did not need the precision of a 16bit int, then they should have put in into a 16bit float.
-This offers a max number of $\pm65504$, rather then the $\pm32768$ a signed int can take.
-Doing so would give a $2\times$ margin of error over the int.
-In addition the conversion should have been protected.
-
-\begin{figure}[H]
-\centering
-\caption{16bit Float, source: \url{https://en.wikipedia.org/wiki/Half-precision_floating-point_format}}
-\label{16}
-\includegraphics{16b.png}
-\end{figure}
-
-However this technique will take a hit on the precision of the number and requires some conversion time.
-Another option is to remove an order of magnitude from the 16bit int, this would give more range and less precision as well.
-Without requiring the extra conversion time.
-It's up to the designer to decide if these are acceptable trade offs.
-\\
-
-\subsection{Testing}
-\label{testing}
-The 501 report \cite{501} lists "There is no evidence that any trajectory data were used to analyse the behaviour of the unprotected variables, and it is even more important to note that it was jointly agreed not to include the Ariane 5 trajectory data in the SRI requirements and specification." \\
-
-This decision wasn't the cause of the failure, however testing like this would have prevented it.
-The last step in designing a Full Fault Tolerant system is extensive testing.
-The telemetry data for the rocket could have been numerically calculated and fed into the system.
-The bug would have made itself obvious and the disaster would have been prevented.
-Burns \cite{BURNS} calls this fault removal and is a major component of fault prevention.
-
-\section{Fault Tolerance}
-% Burns Page 106, should this be Burns? Or Anderson and Lee?
-Burns \cite{BURNS} stated that fault tolerance is  the ability for your system to keep running in the presence of faults.
-This is very challenging in embedded system as there typically can't be any human interaction.
-However when the hardware inevitably fails there should be routines to handle this.
-\\
-
-% The Ariane 5 would have to be ``Full Fault Tolerant'', meaning that in the event of a fault then ship needs to keep flying.
-
-\subsection{Redundancy}
-\label{Red}
-Redundant hardware is used extensively in Full Fault Tolerant system.
-Redundancy provides alternative hardware in the event of a failure of the main hardware.
-Burns \cite{BURNS} wrote about Triple Modular Redundant (TMR) systems where three identical pieces of hardware are running in parallel.
-If the output of one is too far off then the others, it's result is masked out in the computation.
-This same idea can be extended to $N$ pieces of hardware\\
-
-The Ariane 5 had two parallel SRI systems running.
-This means that in the event of a sensor or CPU hardware failure it was possible to switch out to the backup SRI.
-However the failure was not a hardware failure, it was a software failure.
-So both the main and redundant systems failed from the same software failure.
-Therefore redundancy did not save the Ariane 5.\\
-
-Burns \cite{BURNS} spoke of the software equivalent to TMR, so called ``$N$-version'' programming.
-This is a method of software design which which requires $N$-groups of software developers to develop their own codebase, without interaction with each other.
-The software is then run in parallel and takes the same form as TMR.
-\\
-
-However Burns listed three major issues with $N$-Version Programming:
-
-\begin{enumerate}
-\item Poor Initial Specification
-\item Lack of Independence of Design Effort
-\item Inadequate Budget
-\end{enumerate}
-
-Sections \ref{testing} and \ref{execpt} outlined some serious issues with the initial spec, an implementation of $N$-Version programming would have suffered.
-The control system for the Ariane 5 is a complex system, so it would be challenging to prevent people from sharing resources and ideas.
-Finally, would three independent development paths be affordable?
-Typically the most expensive resource of an embeded system is the development time, I don't think it would be reasonable to $3\times$ this.\\
-
-$N-Version$ Programming is a novel approach to the problem and could have been used to avoid the disaster, however the feasibility is still a question. I would recommend approaching the problem
-
-\subsection{Exception Handling}
-\label{execpt}
-The specification for the design stated that during an event of a software exception, the system should indicate the failure on the databus to be stored in EEPROM memory.
-This is followed by a processor shutdown.
-This technique seems reasonable in three situations:
-\begin{enumerate}
-\item Debugging environment.
-\item System with ``N-Version Programming'' where the OBC would detect the exception and another system can take over. (See Section \ref{Red}).
-\item System that catches the exception and can self correct. 
-\end{enumerate}
-This technique in not reasonable for a system that has no backup routine for exceptions.
-The OBC read the software exception bitstream and interpreted as data.
-As the BH routines wasn't even being used at this time of the crash, ignoring the software exception would have been a better response and would have saved the rocket.\\
-
-A somewhat reasonable approach would be to have understood that there was an exception on the databus and extrapolate data.
-This would be an absolute fallback routine that is still very non-ideal, but better than what was implemented.
-
-\subsection{Conversion Protection}
-Baber \cite{SW} has spoken vividly about the conversion. I've written some code below to demonstrate a possible solution that would offer the protection.
-
-\input{$HOME/wlkrUtils/TeX/codeCBW.tex} %For C
-\lstinputlisting{../conv.c}
-After examining the code the flaw becomes obvious, is the event of an overflow the max number will be presented.
-It's possible to make range and precision trade offs, which is discussed more rigorously in Section \ref{SD}.
-If it's not possible to make these trade offs then a simple handshaking routine could be implemented, this is shown in Section \ref{Nconv}\\
-
-\subsection{Handshaking For Conversion}
-It's entirely possible that the team had the processing overhead to implement a handshaking protocol, shown below. 
-\label{Nconv}
-\lstinputlisting{../convN.c}
-This is using a special bit pattern to indicate to the embedded computer that the incoming data will not be a 16bit integer, but a 32 bit integer.
-In addition code should be implemented to ensure that the 64bit number isn't higher than what can fit in the 32bit number.\\
-
-In the system many of the conversions were actually protected.
-However the team claimed that they were not all protected because of the processing overhead.
-I would argue that this is more overhead, but not by much.
-This would do one comparison and would not run the code inside the if statement unless there in in fact an overflow.
-
-\section{Conclusion}
-When developing embedded software the absolute first line of defense is well written code.
-Without a solid foundation there's no amount of testing and fault handling that can save the project.
-That being said, bugs are typically found and fixed during testing afterwords.
-With an extensive test procedure and enough time spent running the system through all scenarios it can become extremely stable.
-The last line of defense is run time software fault handling such as $N$-Version programming and smart use of exceptions.\\
-
-The team that prepared the Ariane 5 for its maiden launch failed at all three of these software development techniques.
-There was poorly written code which made hardware assumptions.
-Following proper code maintenance techniques would have solved the issue from the beginning.
-The team should have started from the drawing board and if any code was to be recycled, it should have been properly studied and understood.
-This would have solved the compatibility issues between the Ariane 4 and Ariane 5.
-There were completely inadequate test procedures.
-A proper testing plan should have been implemented with mock telemetry data.
-If the code maintenance above didn't catch the bug, then testing would have.
-Finally the run time exception handling didn't make any sense.
-There should have been something reasonable implemented in the case of a software exception such as $N$-Version redundancy or data extrapolation.
-
-% Bib
-\begin{thebibliography}{1}
-%\bibitem{501}
-%Prof. J. L. LIONS, "Flight 501 Failure", Report by the Inquiry Board
-%Paris, 19 July 1996
-
-\bibitem{SW}
-Robert L. Baber, "The Ariane 5 explosion as seen by a software engineer", Risks-18.89, 12 March 1997 
-\url{http://www.cas.mcmaster.ca/~baber/TechnicalReports/Ariane5/Ariane5.htm} [Accessed March 12, 2019]
-
-\end{thebibliography}
-\end{document}

### 2019-07-03 report/paper.tex
Added some info about mel spaced filter banks.

@@
 
+\subsection{Mel Spaced Filter Bank}
+Human hearing sensitivity works in a very nonlinear way.
+For a human, there is a stark difference between a signal that is 50hz and a signal that is 150hz.
+However is it almost impossible to notice a difference between a a 10khz signal and a 11khz signal.
+The mel scale has been developed to represent this.
+It it shown in Equation \ref{mel-eq} and is plotted in Figure \ref{mel}.
+
+\begin{equation}
+M(f) = 1125 \cdot log(1+\frac{f}{700})
+\label{mel-eq}
+\end{equation}
+
+\begin{figure}[h]
+\includegraphics[width=9cm]{mel.eps}
+\caption{The Mel Scale}
+\label{mel}
+\end{figure}
+
+To get more adequate feature extraction we are going to use a mel spaced filter bank.
+This effectively maximizes the features extracted as a function of the mel scale.
+The filter array is shown in Figure \ref{FA}.
+In this example we have implemented ten filters for visulizing purposes, however in practice we are going to implement over twenty.
+
+\begin{figure}[h]
+\includegraphics[width=9cm]{FA.eps}
+\caption{The Mel Scale}
+\label{FA}
+\end{figure}
+
 \section{Predicted Difficulties}

### 2019-07-03 report/mel.m

@@ -1,0 +1,11 @@
+f = 0 : 0.1 : 20e3;
+M = 1125*log(1+f./700);
+plot(f, M);
+
+title('The Mel Scale');
+xlabel('f (hz)');
+ylabel('Mel Scale');
+
+set(gca,'FontSize',20);
+
+print('mel.eps');

### 2019-07-03 report/melf.m

@@ -1,0 +1,35 @@
+% Mel Filter bank design
+
+clc ; clear ; close
+
+fmin = 20;
+fmax = 20e3;
+
+mmin = 1125*log(1+fmin/700); % min in mel
+mmax = 1125*log(1+fmax/700); % max in mel
+
+numF = 10; % Ten Filters.
+
+numF = numF + 2;
+
+m = linspace(mmin, mmax, numF);
+f = 700.*(exp(m/1125)-1);
+FB(1) = 0;
+
+figure();
+hold on;
+
+for c = 1:(numF-2)
+  FB(1) = f(c);
+  FB(2) = f(c+1);
+  FB(3) = f(c+2);
+  plot(FB, [0,1,0]);
+end
+
+title('Mel Spaced Filter Bank');
+xlabel('f (hz)');
+ylabel('Mel Scale');
+
+set(gca,'FontSize',20);
+
+print('FA.eps');

### 2019-07-08 report/paper.tex
I've updated the paper
Update layout and reference

@@
 \author{Ryan Walker, Jin Seo}
-
-\IEEEtitleabstractindextext{%
-% \begin{abstract}
-% This paper will outline events of the failed Ariane 5 maiden launch.
-In addition it will describe the failure mode and recommend alternative system design and practices that would have prevented the crash.
-% \end{abstract}
-}
 
@@
 Following the successful deployment we desire to make a bird density heat map of various regions and to undestand how it changes over time.
+Changes here.
 
@@
 
+\section{User Experience}
+Depending on the lifetime of the device, at a regular cadence the user will replace the batteries and SD card.
+As the device will then experience a power cycle, time will be lost.
+The device will then start recording from $t=0$.
+The user is required to keep track of the time of battery replacements.
+Real time is then implemented after the fact.
+
+\section{Signal processing}
+This project is required to utilize several signal processing techniques.
+Some of these techniques could be included filters, discrete Fourier transforms, and so on.
+Filters are necessary to improve the quality of the bird calls and remove unwanted noises.
+In addition, in order to come up with digitized data from acoustic analog signals, other techniques could be applied such as Discrete wavelet transform or mel-frequency cepstrum.
+In this case, MATLAB can be used for simulation.
+
+\subsection{Mel Spaced Filter Bank}
+Human hearing sensitivity works in a very nonlinear way.
+For a human, there is a stark difference between a signal that is 50hz and a signal that is 150hz.
+However is it almost impossible to notice a difference between a a 10khz signal and a 11khz signal.
+The mel scale has been developed to represent this.
+It it shown in Equation \ref{mel-eq} and is plotted in Figure \ref{mel}.
+
+\begin{equation}
+M(f) = 1125 \cdot log(1+\frac{f}{700})
+\label{mel-eq}
+\end{equation}
+
+\begin{figure}[h]
+\includegraphics[width=9cm]{mel.eps}
+\caption{The Mel Scale}
+\label{mel}
+\end{figure}
+
+To get more adequate feature extraction we are going to use a mel spaced filter bank.
+This effectively maximizes the features extracted as a function of the mel scale.
+The filter array is shown in Figure \ref{FA}.
+In this example we have implemented ten filters for visulizing purposes, however in practice we are going to implement over twenty.
+This is Jin's change....
+
+\begin{figure}[h]
+\includegraphics[width=9cm]{FA.eps}
+\caption{The Mel Scale}
+\label{FA}
+\end{figure}
+
+\section{Software}
+In the project we are planning on using Matlab/Simulink for any filter design, C/C++ for embedded code and possibly some python for post processing.
+
 \section{Hardware}
@@
 
-\section{User Experience}
-Depending on the lifetime of the device, at a regular cadence the user will replace the batteries and SD card.
-As the device will then experience a power cycle, time will be lost.
-The device will then start recording from $t=0$.
-The user is required to keep track of the time of battery replacements.
-Real time is then implemented after the fact.
-
-\section{Software}
-In the project we are planning on using Matlab/Simulink for any filter design, C/C++ for embedded code and possibly some python for post processing.
-
-\subsection{Signal processing}
-This project is required to utilize several signal processing techniques.
-Some of these techniques could be included filters, discrete Fourier transforms, and so on.
-Filters are necessary to improve the quality of the bird calls and remove unwanted noises.
-In addition, in order to come up with digitized data from acoustic analog signals, other techniques could be applied such as Discrete wavelet transform or mel-frequency cepstrum.
-In this case, MATLAB can be used for simulation.
-
-\subsection{Mel Spaced Filter Bank}
-Human hearing sensitivity works in a very nonlinear way.
-For a human, there is a stark difference between a signal that is 50hz and a signal that is 150hz.
-However is it almost impossible to notice a difference between a a 10khz signal and a 11khz signal.
-The mel scale has been developed to represent this.
-It it shown in Equation \ref{mel-eq} and is plotted in Figure \ref{mel}.
-
-\begin{equation}
-M(f) = 1125 \cdot log(1+\frac{f}{700})
-\label{mel-eq}
-\end{equation}
-
-\begin{figure}[h]
-\includegraphics[width=9cm]{mel.eps}
-\caption{The Mel Scale}
-\label{mel}
-\end{figure}
-
-To get more adequate feature extraction we are going to use a mel spaced filter bank.
-This effectively maximizes the features extracted as a function of the mel scale.
-The filter array is shown in Figure \ref{FA}.
-In this example we have implemented ten filters for visulizing purposes, however in practice we are going to implement over twenty.
-
-\begin{figure}[h]
-\includegraphics[width=9cm]{FA.eps}
-\caption{The Mel Scale}
-\label{FA}
-\end{figure}
-
 \section{Predicted Difficulties}
@@
 \url{https://atstrack.com/animal-class/avian.aspx} [Accessed June 22, 2019]
+
+\bibitem{BirdCallIdentifier}
+Bird Call Identifier.
+\url{https://web.wpi.edu/Pubs/E-project/Available/E-project-042910-001603/unrestricted/Bird_Call_Identification_MQP_2010.pdf} [Accessed April 29, 2010]
 

### 2019-11-14 NN/audio/src/main.py
Got the

@@ -8,6 +8,7 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 import seaborn as sns
 import tensorflow as tf
+import pdb
 from tqdm import tqdm
 from sklearn.model_selection import KFold, cross_val_score, train_test_split
 
@@ -17,47 +18,23 @@ print(os.listdir("../input"))
 # train = pd.read_json('../input/train.json')
 train = pd.read_json('../../data/1py/data.json')
 
-# Train is a pandas dataframe of the format
-#   >>> train.dtypes
-#   audio_embedding                    object https://github.com/tensorflow/models/tree/master/research/audioset/vggish
-#   end_time_seconds_youtube_clip       int64
-#   is_turkey                           int64
-#   start_time_seconds_youtube_clip     int64
-#   vid_id                             object
-#   dtype: object
-
-
-print(train.shape)
-
 train_train, train_val = train_test_split(train, random_state = 42) # Split into training set and validation set
 
-#xtrain = list()
-#ytrain = list()
-#xval   = list() 
-#yval   = list()
-#
-#for x in train_train:
-#  xtrain.append(train_train['data'][x]['fft'])
-#  ytrain.append(train_train['data'][x]['crow'])
-#
-#for x in train_val:
-#  xval.append(train_val['data'][x]['fft'])
-#  yval.append(train_val['data'][x]['crow'])
-
-xtrain = [k for k in train_train['audio_embedding']]
-ytrain = train_train['is_turkey'].values
-xval = [k for k in train_val['audio_embedding']]
-yval = train_val['is_turkey'].values
+xtrain = [k for k in train_train['fft']]
+ytrain = train_train['crow'].values
+xval = [k for k in train_val['fft']]
+yval = train_val['crow'].values
 
 # Pad the audio features so that all are "10 seconds" long
 
-x_train = tf.keras.preprocessing.sequence.pad_sequences(xtrain, maxlen=10)
-x_val = tf.keras.preprocessing.sequence.pad_sequences(xval, maxlen=10)
+x_train = tf.keras.preprocessing.sequence.pad_sequences(xtrain)
+x_val = tf.keras.preprocessing.sequence.pad_sequences(xval)
 y_train = np.asarray(ytrain)
 y_val = np.asarray(yval)
 
 model = tf.keras.Sequential()
[…]

### 2019-11-15 NN/data/_sandpiper/1py/conv.sh
Data yo

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+for i in *.mp3
+do
+    sox "$i" "$(basename -s .mp3 "$i").wav"
+done

### 2019-11-15 NN/data/1py/conv.sh
Serious re-org

@@ -1,6 +1,0 @@
-#!/bin/bash
-
-for i in *.mp3
-do
-    sox "$i" "waves/$(basename -s .mp3 "$i").wav"
-done

### 2019-11-15 NN/data/_sandpiper/1py/conv.sh

@@ -1,6 +1,0 @@
-#!/bin/bash
-
-for i in *.mp3
-do
-    sox "$i" "$(basename -s .mp3 "$i").wav"
-done

### 2019-11-15 NN/data/fft.py

@@ -6,49 +6,53 @@ from scipy.fftpack import fft, fftfreq, fftshift
 import numpy as np
 import json
 
+folders = ['_crow', '_sandpiper']
+
 folder = '1py/'
 wavfolder = 'waves/'
 fname = '100_crow.wav'
 jsonFname = 'data.json'
-bins = 128 
+bins = 128
 files = [f for f in listdir(folder + wavfolder) if isfile(join(folder + wavfolder, f))]
-dat = {}
-dat['data'] = []
+root = []
 
 for l in range(0, len(files)):
   fname = files[l]
   fs, data = wav.read(folder + wavfolder + fname)   # Open the file
-  if(fs == 44100): 
-    try: 
+  if(fs == 44100):
+    try:
       data = data[:,0]                                # Remove one of the tracks (convert to mono)
     except:
-      continue 
+      continue
     X = fft(data)                                   # Take FFT
     freqs = fftfreq(len(data)) * fs
     freqs = freqs[0:int(len(freqs)/2)]
     X = X[0:int(len(X)/2)]                          # Remove the negative bins
     X = abs(X)                                      # Remove the j shit
     spacing = int(len(X) / bins)                    # Requires spacing
-    
+
     k = list()
     for i in range(0,bins):
       k.append(sum(X[i*spacing:(i+1)*spacing]))
-    
+
     k = k/max(k)
     k = k*256;
     k = [int(i) for i in k]
-    
+
     if "Not" in fname:
       crow = False
     else:
       crow = True
-    
-    dat['data'].append({
+
+    data = []
+    data = k
+
+    root.append({
       'fname': fname,
-      'fft': k,
+      'fft' : data,
       'crow': crow,
-      'fs': fs 
+      'fs': fs
     })
 
 with open(folder + jsonFname, 'w') as outfile:
-  json.dump(dat, outfile)
+  json.dump(root, outfile)

### 2019-11-15 NN/data/splice.py

@@ -1,13 +1,19 @@
+# python splice.py <infolder> <outfolder> <birdName> <StartingFileNumber>
 import os
+import sys
+import pdb
 from os import listdir
 from os.path import isfile, join
 from getch import getch
 
-gCount = 0        # Global Counter
-infolder = "1"    # Folder name to go into
-files = [f for f in listdir(infolder) if isfile(join(infolder, f))]
+gCount = 0                                                         # Global Counter
 
-outfolder = infolder + "py"
+infolder = sys.argv[1]                                             # Folder name to go into
+outfolder = sys.argv[2]
+birdName = sys.argv[3]
+startingPoint = int(sys.argv[4])
+
+files = [f for f in listdir(infolder) if isfile(join(infolder, f))]
 os.system("mkdir " + outfolder)
 
 from pydub import AudioSegment
@@ -16,18 +22,21 @@ from pydub.playback import play
 numFiles = len(files)
 
 for p in range(0, numFiles):
+  print('Opening file: %s' % files[p])
   sound = AudioSegment.from_file(infolder + '/' + files[p])
-  length = int(len(sound)/1000)
-  
-  for k in range(1, length):
-    chunk = sound[(k-1)*1e3:k*1e3]
-    play(chunk)
-    print("y/n")
-    ans = getch()
-    if ans == 'y':
-      out_f = open(outfolder + "/%d_crow.mp3" % gCount, 'wb')
-      chunk.export(out_f, format='mp3')
-    else:
-      out_f = open(outfolder + "/%d_NotCrow.mp3" % gCount, 'wb')
-      chunk.export(out_f, format='mp3')
-    gCount+=1
+  if sound.frame_rate == 44100:  
+    length = int(len(sound)/1000)
+    
+    for k in range(1, length):
+      chunk = sound[(k-1)*1e3:k*1e3]
+      if gCount > startingPoint:
+        play(chunk)
+        print(files[p]+':'+ str(gCount) + ", y/n???")
+        ans = getch()
+        if ans == 'y':
+          out_f = open(outfolder + "/%d_%s.mp3" % (gCount, birdName), 'wb')
+          chunk.export(out_f, format='mp3')
+        else:
+          out_f = open(outfolder + "/%d_Not%s.mp3" % (gCount,birdName), 'wb')
+          chunk.export(out_f, format='mp3')
+      gCount+=1

### 2019-11-15 NN/data/conv.sh
Another big re-org

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+for i in *.mp3
+do
+    sox "$i" "waves/$(basename -s .mp3 "$i").wav"
+done

### 2019-11-15 NN/data/fft.py
Folder rename

@@ -7,6 +7,7 @@ import numpy as np
 import json
 
 folders = ['_crow', '_sandpiper']
+birds   = ['crow', '_sandpiper']
 
 folder = '1py/'
 wavfolder = 'waves/'

### 2019-11-15 NN/data/fft.py
Added on biggo json file

@@ -6,54 +6,52 @@ from scipy.fftpack import fft, fftfreq, fftshift
 import numpy as np
 import json
 
-folders = ['_crow', '_sandpiper']
-birds   = ['crow', '_sandpiper']
-
-folder = '1py/'
-wavfolder = 'waves/'
-fname = '100_crow.wav'
+folders = ['crow', 'sandpiper']
 jsonFname = 'data.json'
 bins = 128
-files = [f for f in listdir(folder + wavfolder) if isfile(join(folder + wavfolder, f))]
 root = []
 
-for l in range(0, len(files)):
-  fname = files[l]
-  fs, data = wav.read(folder + wavfolder + fname)   # Open the file
-  if(fs == 44100):
-    try:
-      data = data[:,0]                                # Remove one of the tracks (convert to mono)
-    except:
-      continue
-    X = fft(data)                                   # Take FFT
-    freqs = fftfreq(len(data)) * fs
-    freqs = freqs[0:int(len(freqs)/2)]
-    X = X[0:int(len(X)/2)]                          # Remove the negative bins
-    X = abs(X)                                      # Remove the j shit
-    spacing = int(len(X) / bins)                    # Requires spacing
+for birds in range(0, len(folders)):
+  files = [f for f in listdir(folders[birds]) if isfile(join(folders[birds], f))]
+  for l in range(0, len(files)):
+    fname = files[l]
+    fs, data = wav.read(folders[birds]+'/'+fname)   # Open the file
+    if(fs == 44100):
+      try:
+        data = data[:,0]                                # Remove one of the tracks (convert to mono)
+      except:
+        continue
+      X = fft(data)                                   # Take FFT
+      freqs = fftfreq(len(data)) * fs
+      freqs = freqs[0:int(len(freqs)/2)]
+      X = X[0:int(len(X)/2)]                          # Remove the negative bins
+      X = abs(X)                                      # Remove the j shit
+      spacing = int(len(X) / bins)                    # Requires spacing
+  
+      k = list()
+      for i in range(0,bins):
+        k.append(sum(X[i*spacing:(i+1)*spacing]))
+  
+      k = k/max(k)
+      k = k*256;
+      k = [int(i) for i in k]
+  
+      if "Not" in fname:
+        isBird = False
+      else:
+        isBird = True
+  
+      data = []
+      data = k
+  
+      root.append({
[…]

### 2019-11-15 NN/audio/init.sh
Got the nn working with the sandpiper data

@@ -8,3 +8,7 @@ chmod +x Anaconda3-2019.10-Linux-x86_64.sh
 source ~/.bashrc                        # Load in the new bashrc
 conda env create --file environment.yml # Create the new env
 conda activate birdear                  # Activate the env
+
+rm Anaconda3-2019.10-Linux-x86_64.sh
+
+# conda env create -f environment.yml

### 2019-11-15 NN/audio/src/main.py

@@ -16,14 +16,14 @@ plt.style.use('fivethirtyeight')
 print(os.listdir("../input"))
 
 # train = pd.read_json('../input/train.json')
-train = pd.read_json('../../data/1py/data.json')
+train = pd.read_json('../../data/data.json')
 
 train_train, train_val = train_test_split(train, random_state = 42) # Split into training set and validation set
 
 xtrain = [k for k in train_train['fft']]
-ytrain = train_train['crow'].values
+ytrain = (train_train['bird'].values == 'crow') & (train_train['isbird'].values)
 xval = [k for k in train_val['fft']]
-yval = train_val['crow'].values
+yval = (train_val['bird'].values == 'crow') & (train_val['isbird'].values) 
 
 # Pad the audio features so that all are "10 seconds" long
 
@@ -31,6 +31,8 @@ x_train = tf.keras.preprocessing.sequence.pad_sequences(xtrain)
 x_val = tf.keras.preprocessing.sequence.pad_sequences(xval)
 y_train = np.asarray(ytrain)
 y_val = np.asarray(yval)
+
+pdb.set_trace()
 
 model = tf.keras.Sequential()
 model.add(tf.keras.layers.Dense(1,input_shape=(128,)))

### 2019-11-17 report/paper.tex
Jin's commit
Worked on the paper.

@@
 %% wlkr
-\documentclass[12pt,journal,compsoc]{IEEEtran}
+\documentclass{article}
 \usepackage[hidelinks]{hyperref}
@@
 \usepackage{tikz}
+\usepackage{pdfpages}
 \usetikzlibrary{positioning}
+\usepackage[margin=1in]{geometry}
+
+\lstset{
+  basicstyle=\scriptsize, %or \small or \footnotesize etc. 
+  commentstyle=\color{red},
+  keywordstyle=\color{blue},
+  showstringspaces=false,
+  breaklines=true,
+  aboveskip=1em,
+  belowskip=1em
+}
 
@@
 \maketitle
-\IEEEpeerreviewmaketitle
 
@@
 \section{Introduction}
-\IEEEPARstart{B}{iologists} have been aiming to understand bird migration for decades.
+Biologists have been aiming to understand bird migration for decades.
 Various custom hardware has been developed for doing this \cite{BirdTracking}, however is it costly an requires capture and release of the bird.
@@
 Following the successful deployment we desire to make a bird density heat map of various regions and to undestand how it changes over time.
-Changes here.
 
 \section{Machine Learning on the Edge}
+\begin{center}
 \begin{tikzpicture}[x=1.3cm, y=1.3cm, >=stealth]
-
 \foreach \m/\l [count=\y] in {1,2,3,missing,4}
@@
 \end{tikzpicture}
+\end{center}
 
-Machine learning and AI have been popular topics in the last couple of years, especially the use of neural nets in classification algorithms.
+Machine learning and AI have been popular topics in the last couple of years, especially the use of neural networks in classification algorithms.
+A neural network is a collection of artificial neurons which can loosely model the neurons in a biological brain.
+They are connected is a way that allows data to propagate through them, each neuron connection (edge) has a weight associated with it, this weight either increases or decreases the strength of the signal at a connection.
+The neurons are arranged in ``layers", the first and last layers are known as the input and output layers respectively, and the middle layers are known at the hidden layers.
+``Training" a neural network is the automated process of adjusting the weights on a sample dataset so that the output gives the requred value for a given input.
+As an example, if were given a problem to classify 1000 pictures into either ``hotdog" or ``not hotdog", where the images were 250 pixels by 250 pixels.
+You would start by:
+
+\begin{itemize}
+\item Manually label a sufficiently large section of the dataset (say 500 pictures)
+\item Introduce 62500 input neurons ($250 \cdot 250$)  
+\item Introduce 1 output neurons (Hotdog or not hotdog)  
+\item Train the network so output says hotdog for all the hotdog pictures, and not hotdog for all the not hotdog pictures
+\item When you then feed in puctures it hasn't seen before, it should classify the remaining data correctly
+\end{itemize}
+
+The accuracy is given by the ratio of correct classifications to incorrect classifications.
 It has not been until recently that these applications have moved into edge compute, meaning all the processing is run on the device as apposed to cloud.
@@
 As this is such a rapidly expanding field there are various software implementations developed for embeded applications \cite{CMSIS} \cite{TF}.
-
-\section{User Experience}
-Depending on the lifetime of the device, at a regular cadence the user will replace the batteries and SD card.
-As the device will then experience a power cycle, time will be lost.
-The device will then start recording from $t=0$.
-The user is required to keep track of the time of battery replacements.
-Real time is then implemented after the fact.
 
@@
 \begin{figure}[h]
+\centering
 \includegraphics[width=9cm]{mel.eps}
@@
 In this example we have implemented ten filters for visulizing purposes, however in practice we are going to implement over twenty.
-This is Jin's change....
 
-\begin{figure}[h]
+\begin{figure}[H]
+\centering
 \includegraphics[width=9cm]{FA.eps}
@@
 
-\section{Software}
-In the project we are planning on using Matlab/Simulink for any filter design, C/C++ for embedded code and possibly some python for post processing.
+\section{Software} In the project we are planning on using Matlab/Simulink for any filter design, C/C++ for embedded code and possibly some python for post processing.
 
@@
 \item{\textbf{Temperature Sensor:} Temperature sensor for recording ambient temperature.}
-\item{\textbf{Additional Sensors???:} Should we put anything else.}
 \end{itemize}
 
-\section{Predicted Difficulties}
-We predict that it will be difficult to gather a a concise and useful dataset to train the neural network.
-However, there are sites that contain vast datasets \cite{Bird}.
-As every circumstance produces different audiable signals, finding proper filters to samples could be a major problem.
-Another major difficulty we predict is implementation of the  Neural Network and how we can apply it to our project, even as we are going to use a library, designing the neural structure for an embedded application might be challenging.
+\section{Data Harvesting and Conditioning}
+In order to train the neural net, we require a great deal of data.
+This was gatherd from a website where bird listeners post sounds of birds \cite{Bird}.
+To this day we have gathered and labeled 589 crow samples, of which roughly half are not crows and 691 sanpiper samples to which roughly half are not sanpipers.
+We did this using the workflow outlined below.
+
+\subsection{Harvesting}
+The script below harvest the data from a given bird.
+\lstinputlisting[language=bash]{../NN/data/fetch.sh}
+
+\subsection{Labeling}
+The labeling process was done manually, it was done with the script below essentually this script will hunt through a file and find sound files, then it will chunk off a 1s piece and play it, if the user presses 'y' then it will save the file down as ``n\_crow" if the user pressed 'n' then it will save the file down a ``n\_Notcrow", where n is an incremented integer.
+\lstinputlisting[language=python]{../NN/data/splice.py}
+
+\subsection{Conditioning}
+Our neural network is deigned to accepts frequency data, not sound data.
+So there is some conditioning required beforehand.
+This script opens up the output files from the last script and preforms a FFT.
+It then reduces the number of bins to something more manageable, right now we are using 128 input neurons, so there are 128 fft bins.
+It then saves all this to a single json file.
+\lstinputlisting[language=python]{../NN/data/fft.py}
+
+\section{The Neural Network}
+We chose Tensorflow for out neural network design because one you train up a model, you have the ability to convert it to a ``tflite" model which can be run on a microcontroller using tensorflow lite.
+We then built the model structure and trained the model.
+Our current model is shown in Figure \ref{model}, it has 128 input neurons on the input layer (for our 128 fft bins) 128 neurons on the one hidden layer and 1 neuron on the out (either crow or not crow).
+The script starts by randomly splitting our dataset into the training and validation set, where the training set it used for the training while the validation set is used for validation after the training is complete, that is to say the network has no knowledge of the validation set when adjusting the weights.
+After this process our network got \textbf{93\% accuracy on the validation set}
+\lstinputlisting[language=python]{../NN/audio/src/main.py}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.3]{../NN/audio/src/model.png}
+\caption{Current Model}
+\label{model}
+\end{figure}
+\newpage
+
+\section{Hardware}
+The first revision of the hardware is finished and we have connected to the microcontroller.
+The physical hardware is shown in Figure \ref{hardware} while the schematic is shown in Figure \ref{sch1} and \ref{sch2}.
+We got a PCB manufactued overseas ans placed all the SMD components onto it.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.15]{../birdEar/img/be.jpg}
+\caption{Hardware}
+\label{hardware}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.7, angle=90, page=1]{../birdEar/plot/birdEar.pdf}
+\caption{Schematic - page 1}
+\label{sch1}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.7, angle=90, page=2]{../birdEar/plot/birdEar.pdf}
+\caption{Schematic - page 2}
+\label{sch2}
+\end{figure}
+
+\section{Next Steps}
+The next steps of the project are...
+
+\begin{itemize}
+\item Implement mel-spaced frequency array.
+\item Increase the number output neurons to 2, so we can classify between crows and sandpipers. 
+\item Add another three types of birds to the dataset.
+\item Research "VGG like feature extraction".
+\item Implement microcontroller firmware for audio sampling.
+\item Implement neural network on the microcontroller.
+\item Deploy the device in the field.
+\end{itemize}
 
@@
 \url{https://atstrack.com/animal-class/avian.aspx} [Accessed June 22, 2019]
-
-\bibitem{BirdCallIdentifier}
-Bird Call Identifier.
-\url{https://web.wpi.edu/Pubs/E-project/Available/E-project-042910-001603/unrestricted/Bird_Call_Identification_MQP_2010.pdf} [Accessed April 29, 2010]
 
@@
 https://www.xeno-canto.org/
+
+\bibitem{BirdCallIdentifier}
+Bird Call Identifier.
+\url{https://web.wpi.edu/Pubs/E-project/Available/E-project-042910-001603/unrestricted/Bird_Call_Identification_MQP_2010.pdf} [Accessed April 29, 2010]
+
 \end{thebibliography}

### 2019-12-06 report/paper.tex
Added some images

@@
 \section{Hardware}
-The first revision of the hardware is finished and we have connected to the microcontroller.
+The first revision of the hardware is finished and we have connected to the microcontroller, we have blinked on of the LEDs but have not implemented any sampling or the neural network.
 The physical hardware is shown in Figure \ref{hardware} while the schematic is shown in Figure \ref{sch1} and \ref{sch2}.
@@
 \centering
-\includegraphics[scale=0.15]{../birdEar/img/be.jpg}
+\includegraphics[scale=0.07]{img/E04A0175.JPG}
 \caption{Hardware}

### 2019-12-10 report/paper.tex
add NN on latex

--- 2019-11-17 report/paper.tex
@@
 \section{The Neural Network}
+\subsection{Background and Exsisting work}
+The idea of Artificial Neural Networks comes from the structure and functions of a human brain. Neurons in human brain that process and transmit information between themselves.
+\begin{figure}[H]
+\includegraphics[scale=0.65]{./NN_general.png}
+\centering
+\caption{Forward\&Backward Neural Network}
+\end{figure}
+
+\subsubsection{Perceptron}
+People used perceptron to model of decision-making. Perceptron works like step function. So, if perceptron is over a certain threshold, it outputs 1 otherwise 0.
+Perceptrons act as a linear decision; however, many reality functions are NOT linear.
+\begin{figure}[H]
+  \centering
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{percept.png}
+    \caption{equation of perceptron.}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{percept2.png}
+    \caption{linearity of perceptron.\protect\cite{Percept}}
+  \end{minipage}
+\end{figure}
+
+\subsubsection{Sigmoid}
+Sigmoid neuron has weights for each input w1,w2,\dots , and overall bias b. 
+It’s $\sigma{(w \cdot x+b)}$ which is called the sigmoid function
+So, the sigmoid can be expressed as:
+
+$\sigma{(z)} =\frac{1}{1+\exp(-z)} \dots (3)$
+
 We chose Tensorflow for out neural network design because one you train up a model, you have the ability to convert it to a ``tflite" model which can be run on a microcontroller using tensorflow lite.
@@
 
+\bibitem{Percept}
+Calculate the Decision Boundary of a Single Perceptron
+https://medium.com/@thomascountz/calculate-the-decision-boundary-of-a-single-perceptron-visualizing-linear-separability-c4d77099ef38
+
+\bibitem{NN}
+Neural Networks and Deep Learning
+http://neuralnetworksanddeeplearning.com/index.html [Accessed Determination Press, June, 2019]
+
 \end{thebibliography}

### 2019-12-11 report/paper.tex
add part of NN on paper
Added BCIT logo and changed the formatting
add NN on paper

@@
 \usepackage[margin=1in]{geometry}
+\usepackage{setspace}
+\doublespacing
 
@@
 }
-
-\begin{document}
-\title{Bird Density, with Edge ML}
-\author{Ryan Walker, Jin Seo}
-
-% make the title area
-\maketitle
 
@@
 }
+
+\begin{document}
+\begin{center}
+\includegraphics[width=6cm]{img/bcit-logo.png} \\
+\huge{Bird Density, with Edge ML} \\
+\large{by} \\
+\large{Ryan Walker and Jin Seo} \\
+\vspace*{\fill}
+\large{\textbf{Bachelors of Technology in Electronics}} \\
+\large{Electrical and Computer Engineering Technology} \\
+\large{British Columbia Institute of Technology} \\
+\large{December - 2019} \\
+\end{center}
+\newpage
+\begin{center}
+\huge{Bird Density, with Edge ML} \\
+\large{by} \\
+
+\large{\textbf{Ryan Walker}} \\
+\large{Diploma in Robotics and Mechatronics, BCIT, 2011} \\
+
+\large{\textbf{Jin Seo}} \\
+\large{Jin put your degree here} \\
+\vfill
+\large{\textbf{Bachelors of Technology in Electronics}} \\
+\large{ELECTRICAL AND COMPUTER ENGINEERING TECHNOLOGY} \\
+
+\vfill
+\large{We accept this report at conforming to the required standard} \\
+\large{.............................................................} \\
+\large{.............................................................} \\
+\large{.............................................................} \\
+\large{.............................................................} \\
+
+\vspace*{\fill}
+\large{\textbf{Bachelors of Technology in Electronics}} \\
+\large{Electrical and Computer Engineering Technology} \\
+\large{British Columbia Institute of Technology} \\
+\large{December - 2019} \\
+\end{center}
+\newpage
+
+\tableofcontents
+\newpage
+
+\listoffigures
+\newpage
 
@@
 \includegraphics[width=9cm]{FA.eps}
-\caption{The Mel Scale}
+\caption{The Mel Space Frequency Array}
 \label{FA}
@@
 \begin{figure}[H]
-\includegraphics[scale=0.65]{./NN_general.png}
+\includegraphics[scale=0.65]{img/NN_general.png}
 \centering
@@
 \subsubsection{Perceptron}
-People used perceptron to model of decision-making. Perceptron works like step function. So, if perceptron is over a certain threshold, it outputs 1 otherwise 0.
+Perceptron that functions as a step function was used to model of decision-making.Thus, if perceptron is over a certain threshold, it outputs 1 otherwise 0.
 Perceptrons act as a linear decision; however, many reality functions are NOT linear.
@@
   \begin{minipage}[b]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{percept.png}
+    \includegraphics[width=\textwidth]{img/percept.png}
     \caption{equation of perceptron.}
@@
   \hfill
-  \begin{minipage}[b]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{percept2.png}
+  \begin{minipage}[b]{0.5\textwidth}
+    \includegraphics[width=\textwidth]{img/percept2.png}
     \caption{linearity of perceptron.\protect\cite{Percept}}
@@
 $\sigma{(z)} =\frac{1}{1+\exp(-z)} \dots (3)$ 
+, $a =\frac{1}{1+\sum_{-j}\exp(-w_jx_j-b)} \dots  (4) $
 
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.75]{img/sig_eqn.png}
+\caption{Activated output neurons with sigmoid.\protect\cite{DL}}
+\label{sig}
+\end{figure}
+
+The Neural network could have multiple layers, so in that case, input of the sigmoid function could be the output of a neuron from previous layer.
+As shown in figure \ref{sig} Different combinations of the activated output neurons fire different neurons in the next layer.
+
+\subsubsection{Learning with Gradient descent}
+
+\includegraphics[scale=0.8]{img/gradient.png}
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/gradient1.png}
+    \caption{Cost function.}
+    \label{cost}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.5\textwidth}
+    \includegraphics[width=\textwidth]{img/gradient2.png}
+    \caption{Gradient descent}
+  \end{minipage}
+\end{figure}
+As shown in the figure \ref{cost},the cost is differences between the expected output and the actual outcome from a neuron .
+Our goal is to find a weights and biases that minimize the cost.
+This is known as gradient descent.Gradient ,in mathematics, means the direction of steepest increase.
+In other words, find a local minimum by training NN.
+
+$a^l =\sigma{(wa^{l-1}+b)} \dots (22)$ ,
+$z^l_k =\sum_{j}-w^l_{kj}a^{l-1}+b^l_k \dots  (23) $\\
+Where $a^l$ is the vector of activations of neuron in the next layer, $a^{l-1}$ is the vector of activations of previous layer of neurons. z is weighted input.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.75]{img/forward.png}
+\caption{Forward propagatoin.}
+\label{forward}
+\end{figure} 
+Unfortunately, when the number of training inputs is very large this can take a long time and learning thus occurs slowly.
+So, stochastic gradient descent can be used to speed up learning.
+The idea is to estimate the gradient C by computing $C_x$ for a small sample of randomly chosen training inputs\protect\cite{NN}Ch2.
+
+\subsubsection{Backward Propagation}
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}[b]{0.8\textwidth}
+    \includegraphics[width=\textwidth]{img/backprop.png}
+    \caption{Backward propagation.\protect\cite{DL}}
+    \label{back}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{img/backprop2.png}
+    \caption{Diagram of Backward propagation.\protect\cite{BK}}
+    \label{back2}
+  \end{minipage}
+\end{figure}
+
+The forward propagation doesn’t give us the best result.
+As mentioned before, we want to find a set of weights and biases to minimize the cost by computing gradient ,and this algorithm is known as the backward propagation.
+As shown the figures \ref{back} and \ref{back2}, partial derivative and chain rule drive the network backward.
+
+\subsubsection{Cross-entropy Cost function}
+As shown in the sigmoid \ref{sig}, if the neuron's output get close to 1, the curve gets flat, so $\sigma^{'}(z)$ get very small (slow down).
+Cross-entropy avoids the problem of slowdown by vanishing the term $\sigma^{'}(z)$
+
+\includegraphics[scale=0.8]{img/cross.png}, 
+\includegraphics[scale=0.6]{img/cross1.png}
+
+the term $\sigma^{'}(z)$ disappears in the equation (61) after derivative of the cost with respect to weight.
+
+\subsubsection{Training,Validation,and Early stopping}
+In training process, separate training and validating data so as to evaluate NN model.
+In a training process, we define the number of epochs that iterates forward and backward propagation.
+Perhaps people think of increasing the number of epochs leads us to achieve the better result.
+However, there is a problem called over fitting.
+Over fitting (Over training) is a situation where the cost of training data decreases but the accuracy of validating data gets saturated.
+Once the classification accuracy on the validation data has saturated, we stop training (Early stopping).
+\begin{figure}[H]
+  \centering
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/over.png}
+    \caption{Decreasement of cost as epochs increases in training data.\protect\cite{NN}}
+    \label{over}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/over1.png}
+    \caption{Saturation of accuracy in validating data.\protect\cite{NN}}
+    \label{over1}
+  \end{minipage}
+\end{figure}
+
+To avoid overfitting.
+Then save the hyper-parameters (\# of epochs, learning rate,...).
+The learning rate is how fast the network learns; however, the network learning fast could cause over-shooting in the gradient descent.
+Repeating training by modifying the NN model and parameters until getting good accuracy.
+
+\subsubsection{Other NN model - Convolutional NN}
+another type of NN called Convolutional NN. $a^l=\sigma{(a^{l-1}*w+b)}$and $\ast$ is called a convolution operation.
+
+\includegraphics[scale=0.8]{img/con_eqn.png}
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}[b]{0.6\textwidth}
+    \includegraphics[width=\textwidth]{img/convol.png}
+    \caption{Break down maps and features in CNN.\protect\cite{CON}}
+    \label{over}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.6\textwidth}
+    \includegraphics[width=\textwidth]{img/convol1.png}
+    \caption{Parameters reducing by CNN.\protect\cite{CON}}
+    \label{over1}
+  \end{minipage}
+\end{figure}
+
+As you see in the figure \ref{over1}, the outcome of filtering is compressed by the filtering steps. In other words each element doesn’t have own weights and biases, they share weights and biases!.
+
+Filtering steps {\protect\cite{CON}}:
+\begin{itemize}
+\item Line up the feature and map
+\item Multiply each element of map by corresponding feature element
+\item Add them up
+\item Divide by the total number of elements in the feature
+\end{itemize}
+A big advantage of sharing weights and biases is that it greatly reduces the number of parameters involved in a convolutional network.
+What is good about reducing the number of parameters? First,it speeds up the computation. Second,it gives us better accuracy.An example of using image pixels has been used to explain CNN.
+What’s about our case?
+Our input getting into the NN is MFCC(Mel Frequency Cepstral Coefficients).
+As we visualize the data of MFCC, it does not look quite different from the image data!.A spectrogram is a visual representation of the spectrum of frequencies of a signal as it varies with time.
+
+\begin{figure}[H]
+\includegraphics[scale=0.25]{img/MFCC.png}
+\caption{MFCC \protect\cite{AU}}
+\label{mfcc}
+\end{figure}
+
+\subsubsection{Max-pooling}
+In max pooling, a pooling unit simply outputs the maximum activation in the N×N input region.
+Pooling steps {\protect\cite{CON}}:
+\begin{itemize}
+\item Pick window size( 2 or 3)
+\item Pick stride size (usually 2)
+\item Walk the window across the filtered outputs (CNN)
+\item From each window, take the maximum value
+\end{itemize}
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}[b]{0.6\textwidth}
+    \includegraphics[width=\textwidth]{img/max.png}
+    \caption{Max-pooling.\protect\cite{CON}}
+    \label{max}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.6\textwidth}
+    \includegraphics[width=\textwidth]{img/max1.png}
+    \caption{Model of CNN and Max-pooling.\protect\cite{DL}}
+    \label{max1}
+  \end{minipage}
+\end{figure}
+
+\hfill \break A big benefit is that this helps reduce the number of parameters needed in later layers as shown in the figure \ref{max},
+
+\subsubsection{Other activation function - Rectified Linear Units (ReLU)}
+The concept and example of Neural Network that we have dealt with so far use the one activation function which is the sigmoid.
+There are different type of activation functions and one of those is Rectified Linear Units.
+\begin{figure}[H]
+\includegraphics[scale=0.8]{img/relu.png}
+\caption{ReLu \protect\cite{CON}}
+\label{rele}
+\end{figure}
+\hfill \break ReLu nomalizes the output of NN by tweaking each of values just a bit ex) change every negative values to zero \cite{DL}.
+People found that networks based on rectified linear units consistently outperformed than networks based on sigmoid activation functions.However,it has shown that ReLu is empirically better( Not mathematically proved).
+
+\subsection{Put all together}
+
+\begin{figure}[H]
+\includegraphics[scale=0.7]{img/all.png}
+\caption{Model of Birdear}
+\label{mod}
+\end{figure}
+As shown in the figure \ref{mod}, the signal of bird chirpping goes into our the signal process.
+It ends up composing the MFCC, which is the input of NN.
+In the end, the NN distingusih what kind of bird is.
+
+\subsection{Software Appication(PC) and Experiments}
 We chose Tensorflow for out neural network design because one you train up a model, you have the ability to convert it to a ``tflite" model which can be run on a microcontroller using tensorflow lite.
@@
 After this process our network got \textbf{93\% accuracy on the validation set}
+
+\subsubsection{Testing codes(Python)}
+Following codes represent NN model that has a couple of CNN and a Max-pooling:
+\lstinputlisting[language=python]{../NN/audio/src/train.py}
+\hfill \break
+Another simple NN:
 \lstinputlisting[language=python]{../NN/audio/src/main.py}
@@
 \bibitem{Percept}
-Calculate the Decision Boundary of a Single Perceptron
-https://medium.com/@thomascountz/calculate-the-decision-boundary-of-a-single-perceptron-visualizing-linear-separability-c4d77099ef38
+Calculate the Decision Boundary of a Single Perceptron,
+Tomas Countz,
+\url{https://medium.com/@thomascountz/calculate-the-decision-boundary-of-a-single-perceptron-visualizing-linear-separability-c4d77099ef38}, Apri 2018
 
 \bibitem{NN}
-Neural Networks and Deep Learning
-http://neuralnetworksanddeeplearning.com/index.html [Accessed Determination Press, June, 2019]
+Neural Networks and Deep Learning,
+Michael Nielsen,
+\url{http://neuralnetworksanddeeplearning.com/index.html} [Accessed Determination Press, June, 2019]
+
+\bibitem{DL}
+Deep learning,
+\url{https://www.youtube.com/watch?v=aircAruvnKk&list=PLZHQObOWTQDNU6R1_67000Dx_ZCJB-3pi&index=1},
+2017
+
+\bibitem{BK}
+A beginner’s guide to deriving and implementing backpropagation,
+Pranav Budhwant,
+\url{https://medium.com/binaryandmore/beginners-guide-to-deriving-and-implementing-backpropagation-e3c1a5a1e536}
+
+\bibitem{CON}
+How Convolutional Neural Networks work,
+Brandon Rohrer,
+\url{https://www.youtube.com/watch?v=FmpDIaiMIeA&t=243s},
+2016
+
+\bibitem{AU}
+Speech Processing for Machine Learning: Filter banks, Mel-Frequency Cepstral Coefficients (MFCCs) and What's In-Between,
+Haytham Fayek,
+\url{https://haythamfayek.com/2016/04/21/speech-processing-for-machine-learning.html},
+April, 2016
+
+\bibitem{MF}
+Mel Frequency Cepstral Coefficient (MFCC) tutorial,
+James Lyons,
+\url{http://practicalcryptography.com/miscellaneous/machine-learning/guide-mel-frequency-cepstral-coefficients-mfccs/},
 

### 2019-12-12 report/paper.tex
Complete NN and Conclusion, Appendix, etc on paper

@@
 
-\large{\textbf{Jin Seo}} \\
-\large{Jin put your degree here} \\
+\large{\textbf{Jin Won Seo}} \\
+\large{Diploma of Computer Control in ECET, BCIT, 2011} \\
 \vfill
@@
   \centering
-  \begin{minipage}[b]{0.4\textwidth}
+  \begin{minipage}[b]{0.3\textwidth}
     \includegraphics[width=\textwidth]{img/percept.png}
@@
 $\sigma{(z)} =\frac{1}{1+\exp(-z)} \dots (3)$ 
-, $a =\frac{1}{1+\sum_{-j}\exp(-w_jx_j-b)} \dots  (4) $
+, $a =\frac{1}{1+\sum_{-j}\exp(-w_jx_j-b)} \dots  (4)\cite{NN}$
 
@@
 In other words, find a local minimum by training NN.
-
-$a^l =\sigma{(wa^{l-1}+b)} \dots (22)$ ,
-$z^l_k =\sum_{j}-w^l_{kj}a^{l-1}+b^l_k \dots  (23) $\\
+$\\a^l =\sigma{(wa^{l-1}+b)} \dots (22)\cite{NN} ,\\
+z^l_k =\sum_{j}-w^l_{kj}a^{l-1}+b^l_k \dots  (23)\cite{NN}\\$
 Where $a^l$ is the vector of activations of neuron in the next layer, $a^{l-1}$ is the vector of activations of previous layer of neurons. z is weighted input.
@@
 As shown the figures \ref{back} and \ref{back2}, partial derivative and chain rule drive the network backward.
+Small changes of weights change the cost; therefore, the chain rule can be applied to explain this phenomenon.
+$\frac{\partial C_0}{\partial w^L} = \frac{\partial z^L}{\partial w^L}\frac{\partial a^L}{\partial w^L}\frac{\partial C_0}{\partial a^L}\dots (7.1)\\$
+As we take a derivative of each terms.
+$\\ \frac{\partial z^L}{\partial w^L}=a^{L-1} \dots (7.2)
+\\ \frac{\partial a^L}{\partial z^L}=\sigma^{'}(z^L) \dots (7.3)
+\\ \frac{\partial C_0}{\partial w^L} = 2(a^L - y) \dots (7.4)$
+\\As you see the term $,\frac{\partial z^L}{\partial w^L}=a^{L-1},$ partial derivative gives us the activated output neuron in the previous layer.
+This doesn't explain whole equations of backward propagation, but it gives us an idea how the backward propagation works.
 
@@
 \subsubsection{Other NN model - Convolutional NN}
-another type of NN called Convolutional NN. $a^l=\sigma{(a^{l-1}*w+b)}$and $\ast$ is called a convolution operation.
 
-\includegraphics[scale=0.8]{img/con_eqn.png}
-
+Another type of NN called Convolutional NN. $a^l=\sigma{(a^{l-1}*w+b)}$and $\ast$ is called a convolution operation.
+\\Activation of CNN = $\sigma{(b+\sum_{l=0}^{n}\sum_{m=0}^{n}-w_{l,m}a_{j+l,k+l})} \dots  (125)\cite{NN}$
 \begin{figure}[H]
@@
 \end{figure}
-As shown in the figure \ref{mod}, the signal of bird chirpping goes into our the signal process.
+As shown in the figure \ref{mod}, signals of bird chirpping go into our the signal process.
 It ends up composing the MFCC, which is the input of NN.
@@
 After this process our network got \textbf{93\% accuracy on the validation set}
-
-\subsubsection{Testing codes(Python)}
-Following codes represent NN model that has a couple of CNN and a Max-pooling:
-\lstinputlisting[language=python]{../NN/audio/src/train.py}
-\hfill \break
-Another simple NN:
-\lstinputlisting[language=python]{../NN/audio/src/main.py}
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.3]{../NN/audio/src/model.png}
-\caption{Current Model}
-\label{model}
-\end{figure}
-\newpage
 
@@
 \end{figure}
+
+\section{Conclusion}
+As mentioned in the section "Data Harvesting and Conditioning",
+we have gathered and labeled 589 crow samples, of which roughly half are not crows and 691 sanpiper samples to which roughly half are not sanpipers.
+Those samples are stored as WAV formats so that the modules of signal process can extract MFCC at the end of the process.
+Since MCU has a limited resources, it is necessary to compress digital data as much as possible.
+However, just compressing data into MCU is not our goal.
+The idea is how much signal data could be compressed without lossing discrete digital data in low frequency range.
+In other words, the compressed data still should represent the features of the original signal.\\
+Sound of birds' chirpping keeps changing abrumptly in a short time; therefore, it may not make so much sense to take FFT(Fast Fourier Transform) in the whole signal.
+In this case, STFT (Short time Fourier Transform) could be better solution.
+In addition, applying DCT(Discrete Cosine Transform) could help to decompose the overlapped signals from the STFT and compress data.
+The reason to come up with MFCC(Mel Frequency Cepstral Coefficients) is that first, it allows to shrink down the size of inputs.
+Second, it filters out signals in high frequency range, and focuses on the human perceivable hearing range (low frequency range).
+Our goal is not making some high quality audio sounds; rather, we are processing the signals such that NN can be discriminative enough between the birds' chirpping.Perhaps, the testing result of NN,around 93\%, still looks promissing with our testing samples; however, in order to solidify the model of NN, more diverse data could be required.
+The more varioud data we have, the better accuracy the NN performs.
+But, getting more data results in expensive cost since it takes a time and efforts.\\
+Unlike the other typical embedded system projects that are directly designed and implemented on MCU.
+Our project requires pre-steps before attacking MCU, which is to verify the model in the environments of PC.
+Although PCs and embedded systems are quite different in a matter of archetecture and resources, it seems to be inevitable to work on PC first.
+The reasons behind is that first, Machine Learing is still a new area for the embedded systems.
+Another reason is lots of resources of NN work based on PC, so without getting through the test based on PC, it is hard to estimate that the model of NN works on the embedded systems.\\
+From this project, we have been exposed ourselves to a new rising technical area, Neural Network, to see if it is possible to apply this to the embedded systems.
+Our answer for this is of course it is possible, and there seems to be area where electrical and electronics engineering could adapt the new rising technique such as signal process, power system, tele-communication, and so on.
+Moreover, we have learned techniques of the signal process that handle analogue to digital signals demanding compressing data for the embedded systems.
+\newpage
 
@@
 James Lyons,
-\url{http://practicalcryptography.com/miscellaneous/machine-learning/guide-mel-frequency-cepstral-coefficients-mfccs/},
+\url{http://practicalcryptography.com/miscellaneous/machine-learning/guide-mel-frequency-cepstral-coefficients-mfccs/}
 
 \end{thebibliography}
+\newpage
+
+\appendix
+\section{Appendix A}
+\subsection{Testing NN codes(Python)}
+
+Following codes represent NN model that has a couple of CNN and a Max-pooling:
+\lstinputlisting[language=python]{../NN/audio/src/train.py}
+\hfill \break
+Another simple NN:
+\lstinputlisting[language=python]{../NN/audio/src/main.py}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.3]{../NN/audio/src/model.png}
+\caption{Current Model}
+\label{model}
+\end{figure}
+\newpage
+\subsection{Testing Signal Process codes(Python)}
+Read .wav that recorded bird chirpping, then feed it into the signal process that outputs MFCC(Mel Frequency Cepstral Coefficients)
+
+\lstinputlisting[language=python]{../NN/data/SignalProcess/stft_mel.py}
+
+\section{Appendix B}
+\subsection{List of abbreviations and acronyms}
+
+STFT = Short Time Fourier Transform\\
+FFT = Fast Fourier Transform\\
+DCT = Discrete Cosine Transform\\
+MFCC = Mel Frequency Cepstral Coefficients\\
+MCU = Micro Controller\\
+PC = Personal Computer\\
 
